State of the Practice for Lattice Boltzmann Method Software

### 2022-02-26 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
First draft of Section 9.1 (Design for change)

	\item[Reproducibility] Software achieves reproducibility if another
	developer can take the requirements documentation and re-obtain the same
	software artifacts \citep{BenureauAndRougier2017}. This includes the output
	of the software, where the scientific results are compared between software
	implementations, or between software implementations and manually calculated
	results. We measured reproducibility qualitatively by asking developers if
	they have any concern that their computational results won't be
	reproducible, and if they have taken any steps to ensure reproducibility.

@@

Increasing source code modularity, reducing duplicate information,
and improving abstraction by developing well defined interfaces. This was noted
by the developers of ESPResSo and pyLBM.  A developer of pyLBM mentioned that
the geometries and models of their system had been ``decoupled'', using
abstraction and modularization of the source code, to make it ``very easy to add
[new] features''.  \wss{Highlights the advantage of design.  Other projects can
look to these examples.  Maybe cite the Anshu paper about Flash? Dune? and point
to the advantages of design.} Addresses maintainability.  The developer of TCLB
mentioned that their package had two sets of code, for executing the models on
the CPU and GPU, and that maintenance was decreased by introducing macros, a
practice which became a common part of the development process.

A high degree of code modularity and abstraction was noted by developers as a
measure to ensure the ease of future changes. This can be ensured by separating
components and hiding information behind well defined interfaces. The developer
of ESPResSo also noted that some of the code base was transitioned from C to
C++, which could ease modifiability of that software package. The developer of
TCLB noted that their software package was designed to allow for the addition of
some LBM features, but changes to major aspects of the system would be
difficult. For example, ``implementing a new model will be an easy
contribution'', but changes to the ``Cartesian mesh...will be a nightmare''.
\wss{example suggests identifying likely changes.}

Modularize the source code, separate components, hide information behind well
defined interfaces. This is suggested in SCS literature, and in developer
comments.  \wss{one of Wilson's best practices, although Wilson (like so many
others) is silent on what criteria to use for modularization.}

Modifiability. Consider flexibility of data structures and data
storage in the design stage. The package pyLBM redeveloped data structures
to ease future changes.

Contrast with SE literature.  Can point to references, starting with Parnas, on
the importance of modularization.  Also they seem to have a basis for the
modularization that matches with Parnas - likely changes.

@@

>>>>

	\item[Reproducibility] Software achieves reproducibility if another
	developer can take the original code and input data, run it, and obtain the
	original output \citep{BenureauAndRougier2017}. We measured reproducibility
	qualitatively by asking developers if they have any concern that their
	computational results won't be reproducible, and if they have taken any
	steps to ensure reproducibility.

@@

The top LBM developers show how modularization can be used to design research
software for future change.  Although the advice to modularize research software
to handle complexity is common \citep{WilsonEtAl2014, StewartEtAl2017,
Storer2017}, specific guidelines on how to divide the software into modules is
less prevalent.  Not every decomposition is a good design for supporting change,
as shown by \citet{Parns1972a}.  A design with low cohesion and
high coupling \citep[p.\ 48]{GhezziEtAl2003} will make change difficult.
Especially in research software, where change is inevitable, designers need to
produce a modularization that supports change.

LBM developers highlighted in our interviews cases where their modularizations
anticipated future changes.  The developer of pyLBM mentioned that the
geometries and models of their system had been ``decoupled'', using abstraction
and modularization of the source code, to make it ``very easy to add [new]
features''.  The pyLBM design allows for independent changes to the geometry and
the model.  We also learned that the package pyLBM redeveloped data structures
to ease future changes. The developer of TCLB noted that their software package
was designed to allow for the addition of some LBM features, but changes to
major aspects of the system would be difficult. For example, ``implementing a
new model will be an easy contribution'', but changes to the ``Cartesian mesh
... will be a nightmare''.  The design of TCLB highlights that not every conceivable change needs to be supported, only the likely changes.  Likely changes.

As the LBM developers illustrate, design for changes is accomplished by first
identifying likely changes, either implicitly or explicitly, and second by
hiding those likely changes behind well-defined module interfaces.

@@

\subsection{Reproducibility}

- record versions of everything
- Docker, Virtual machines
